Switchable substrates and thin film morphologies

### 2023-04-21 Reply_RefereeA.tex

\section*{Reply to Referee A}
We thank the Referee for his/her report on our manuscript. 
We are glad to read that he/she finds it of potential interest and in principle 
suitable for publication, if a thorough revision is performed. 
We substantially improved and revised the paper for the resubmission and hope that we were able to clarify the obscure points, following all the Referee's remarks.
Hereafter, we provide answers to the various questions/comments raised. All changes in the revised version are reported in red color.

\begin{itemize}

\item[ \textbf{\underline{Comment 1.}}]
{ 
I am unable to follow most of the arguments made with respect to the disjoining pressure. 
For me, one needs to integrate the disjoining pressure to arrive at an effective interface potential, and 
I do not really know how to go from the effective interface potential to the surface tension. 
The equation (3) therefore merits much more discussion.
}

\item[ \textbf{{Answer}}]
{
We did not give, in fact, any particular argument with respect to the disjoining pressure, since 
we assumed that it is a very well established concept employed in any theoretical and computational study of thin liquid film hydrodynamics in the lubrication approximation framework.
We also assumed that an equation like Eq.(3) (now Eq.(2)), which is almost thirty years old \cite{SCHWARTZ1998173,Mitlin94}, did not deserve a lengthy discussion.
However, our assumptions were probably erroneous and, therefore, we give, here, some more explanation. 
Let us, first, make a step back and underline a fundamental difference between the physics of {\it one free interface} and that of a {\it film}:
the latter is a layer of liquid confined between {\it two} interfaces. When the two interfaces are brought close enough to each other (the layer is thin), they interact.
The disjoining pressure stems from this interaction. Thermodynamically, it represents minus the derivative of the free energy difference, per unit surface, between 
the film and the bulk of the same liquid, or, equivalently, the pressure difference inside and outside the film 
\cite{Deryaguin1940,DeryaguinChuraev1978}.
In the case of wetting, there can be three types of interfaces: liquid/air, liquid/solid (the substrate) and, 
if the substrate is only partially wetted, solid/air, thus giving rise to three-phase contact lines. 
The usually called {\it surface tension} is the energy cost of forming a liquid/air interface of unit area and its only knowledge is, therefore, not enough in our case. 
The Young's condition \cite{Young1805,DeGennes1985} connects the surface tension with the analogous energies per unit
area associated with the solid/liquid ($\gamma_{\text{sl}}$) and solid/gas ($\gamma_{\text{sg}}$) interface as
\begin{equation}
\cos \theta = \frac{\gamma_{\text{sg}}-\gamma_{\text{sl}}}{\gamma},
\end{equation}
where $\theta$ is the contact angle.
So, there is no need (and no way) to go from the disjoining pressure (or the interface potential) to the surface tension, since the two quantities carry complementary physical information,
and both are required to determine the dynamics of dewetting.
We have included these comments in a section (called 'Disjoining pressure') 
in the new Supplemental Material.
Of course, on the other hand, the disjoining pressure depends also on the surface tension, which appears, in fact, in its expression (2). 
We think, though, that part of the confusion might be due to the fact that we do not define the symbol of the surface tension at its first occurrence, but only later. 
We have now fixed this and we apologize for it. 
}

@@

\textcolor{red}{Lengths and time scales will be expressed, respectively, in units of the mean film height, $h_0$ (which is constant in time, due to mass conservation), and of $t_0 = \frac{3\mu}{\gamma h_0^3 q_0^4}$, the inverse growth rate of the most unstable mode, whose wavenumber is $q_0$, of a spinodally dewetting film. On a uniform substrate, with constant contact angle $\theta^{(u)}$, the wavenumber reads 
$(q^{(u)}_0)^2 = h_{\ast}^{-2}(1-\cos \theta^{(u)})f^{\prime}(h_0/h_{\ast})$ \cite{Mecke_2005,PhysRevE.100.023108}. In our patterned case, we define 
$q_0^2=h_{\ast}^{-2}(1-\cos\theta_0)f^{\prime}(h_0/h_{\ast})$.
Correspondingly, we choose as a velocity scale $v_0 = \lambda_s/t_0$, where $\lambda_s = 2\pi/q_0$.}

}

\item[ \textbf{\underline{Comment 3.}}]
{ 
Incidentally, for a 'normal' fluid such as water, the capillary velocity is $\sim 70\mbox{m}/\mbox{s}$, 
so that very, very rapidly the fluid inertia becomes important. 
This is indeed what one usually observes for the formation or coalescence of drops, so that at the very least this should be discussed and the Reynolds number calculated.
}

\item[ \textbf{{Answer}}]
{
Such a high value corresponds to the ratio of (surface tension)/(dynamic viscosity) for water, i.e. $v = \gamma/\eta \approx 70 \mbox{m}/\mbox{s}$.
This expression, though, is only representative of the capillary velocity, again, for {\it free surface flows}, as, e.g., in the coalescence of 
two droplets in the bulk of another fluid. 
In the case wetting/dewetting, 
the phenomenology is complicated by the liquid/substrate (and vapour/substrate) interaction. 
In such a situation the relevant velocity scale is set by the characteristic velocity of contact line 
motion which can be estimated from a viscocapillary balance as 
$\sim \varepsilon^3 \gamma/\eta$, where $\varepsilon = h_0/L \ll 1$ is the film thickness parameter 
\cite{RevModPhys.69.931,doi:10.1146/annurev-fluid-011212-140734}. 
The typical velocities in dewetting problems are, then, much smaller than the ratio $\gamma/\eta$.
Moreover, in many actual applications, as well as in the majority of experiments, the 
working fluids are in general much more viscous than water (e.g., polymeric liquids), thus making
the characteristic velocity even smaller.
For the dewetting of a millimetric film of micrometric thickness reatracting into a single droplet, for instance, experiments have measured that 
the typical velocity is $\sim 1 \mbox{mm}/\mbox{s}$, which gives a Reynolds number $Re \approx 10^{-3}$ \cite{Edwards2016}. 
In our problem, there are two relevant characteristic velocities: the retraction speed, 
$U_{\Theta}$, and 
the pattern speed, $v_{\theta}$. When their ratio $\Gamma = v_{\theta}/U_{\Theta} < 1$, the retraction speed dominates and fixes the Reynolds number which can be, then, evaluated as 
$Re = U_{\Theta}h_0/\nu$ \cite{RevModPhys.69.931}, whereas for $\Gamma > 1$, $Re = v_{\theta} h_0/\nu$. 
In any case, the Reynolds number stays always below $\approx 0.2$.
We have added the following comment in the Supplemental Material in the section on 'Parameter values and comparison with experiments':\\

@@

\item[ \textbf{\underline{Comment 4.}}]
{ 
In addition, there is a lot of old literature that relates the dewetting to the disjoining pressure, and shows that the nucleation of a hole in a wetting film is much more difficult than the nucleation of such a wetting film; there are experiments on this by Rutledge and Taborek, theory by Schick and Taborek and Bausch, Blossey and Indekeu.
}

\item[ \textbf{{Answer}}]
{
We are well aware of the literature quoted by the Referee, but these are works 
about {\it prewetting}, namely the phenomenon of {\it nucleation} (condensation)
of films/droplets on solid surfaces confining a critical liquid-vapour mixture
{\it as a function of temperature}. So, we are a bit puzzled about what such literature may have to do with the {\it dewetting} of a film on a chemically patterned substrate. Here, in fact, we are modelling a thin liquid film which is thermodaynamically stable against evaporation (i.e. one can consider that 
we are at a - fixed - temperature below the critical one).
}

\item[ \textbf{\underline{Comment 5.}}]
{ 
The explanations also lack a discussion on time, length, and energy scales. 
I have no clue what the values mean ``the numerical values, in LB units, are set to $\gamma= 0.01$ and $\mu = 1/6$''?. 
}

\item[ \textbf{{Answer}}]
{
Length and time scales, as discussed in Comment 2 and as mentioned also in the original version of the manuscript, 
are made dimensionless with the mean film height and the characteristic time $t_0$. 
These two quantities 
have a clear physical meaning and can, therefore, be used to compare with 
experimental setups. In particular, the expression of $t_0$, through its dependence 
on $q_0$ (the wavenumber of the most unstable mode) contains the details 
of the disjoining pressure, such as the Hamaker constant $A_H$. For instance, let us consider the problem studied in \cite{becker2003complex}, the dewetting of a polystyrene film of thickness $h_0 \approx 4 \, \text{nm}$ deposited on a silicon 
dioxide substrate. The Hamaker constant for this system is 
$A_H  = 2.2 \times 10^{-20} \, \text{J}$ and the surface tension is 
$\gamma = 0.03 \, \text{N}/\text{m}$, such that, 
from equation (3) of \cite{becker2003complex} and 
from the definition of $q_0^2 = \frac{1}{2\gamma} \Pi^{\prime}(h_0)$, 
we get (neglecting the short range part of the interface potential)
$$
q_0^2 \approx \frac{A_H}{4\pi \gamma h_0^4} \approx 2.5 \times 10^{-4} \, \text{nm}^{-2}.
$$
The characteristic time of the growth of the instability, then, reads \cite{PhysRevLett.99.114503} (the dynamic viscosity of polystyrene is 
$\mu = 1.2 \times 10^4 \, \text{Pa} \cdot \text{s}$)
$$
t_0 = \frac{3\mu}{\gamma h_0^3 q_0^4} \approx 300 \, \text{s}
$$
and it can be seen that indeed film rupture occurs within time 
scales of the order of $t_0$ \cite{becker2003complex}.
We have reported these comments on comparison with actual experimental 
values in the new Supplemental Material as follows:\\
\\
\textcolor{red}{To give a flavour of how to compare with actual experimental systems, let us notice
first that the expression of $t_0$ depends on the disjoining pressure (through its dependence 
 $q_0 = \sqrt{\frac{1}{2\gamma} \Pi^{\prime}(h_0)}$) and contains, therefore, energetic details, namely 
 the Hamaker constant $A_H$. Consider, for instance, the problem studied in \cite{becker2003complex}: the dewetting of a polystyrene film of thickness $h_0 \approx 4 \, \text{nm}$ deposited on a silicon 
dioxide substrate, where $A_H  = 2.2 \times 10^{-20} \, \text{J}$ and 
$\gamma = 0.03 \, \text{N}/\text{m}$. Neglecting the short range part of the interface potential
(such that $\Pi(h)= - \frac{A_H}{6\pi h^3}$)~\cite{Mecke_2005,becker2003complex}, one gets 
$q_0^2 \approx \frac{A_H}{4\pi \gamma h_0^4} \approx 2.5 \times 10^{-4} \, \text{nm}^{-2}$.
The characteristic time of the growth of the instability for these experiments, then, reads \cite{PhysRevLett.99.114503} (the dynamic viscosity of polystyrene is 
$\mu = 1.2 \times 10^4 \, \text{Pa} \cdot \text{s}$) $t_0 = \frac{3\mu}{\gamma h_0^3 q_0^4} \approx 300 \, \text{s}$ and it can be seen that indeed film rupture occurs within time 
scales of the order of $t_0$ \cite{becker2003complex}, as in our 
simulation for the largest pattern wavelength, $\lambda = L$ (see also our previous work
with homogeneous substrates~\cite{PhysRevE.104.034801}).} 
\\

### 2023-04-21 Old_version/main-veryold.tex
New replies, thanks for nothing ref A

%\title{Switchable substrates and thin film morphologies}
\title{Spatiotemporal varying contact angles and thin film morphologies}

\author{S. Zitz}
\email{s.zitz@fz-juelich.de}
 \affiliation{Helmholtz Institute Erlangen-N\"urnberg for Renewable Energy,\\
  Forschungszentrum J\"ulich,\\
  F\"urther Strasse 248, 90429 N\"urnberg, Germany}%
  \affiliation{Department of Chemical and Biological Engineering, Friedrich-Alexander-Universit\"at Erlangen-N\"urnberg, F\"{u}rther Stra{\ss}e 248, 90429 N\"{u}rnberg, Germany}
\author{A. Scagliarini}%
\email{andrea.scagliarini@cnr.it}
 \affiliation{Institute for Applied Mathematics "M. Picone" (IAC), \\
Consiglio Nazionale delle Ricerche (CNR),\\
Via dei Taurini 19, 00185 Rome, Italy}%
\affiliation{INFN, sezione Roma ``Tor Vergata'', via della Ricerca Scientifica 1, 00133 Rome, Italy}
\author{J. Harting}
\email{j.harting@fz-juelich.de}
 \affiliation{Helmholtz Institute Erlangen-N\"urnberg for Renewable Energy,\\
  Forschungszentrum J\"ulich,\\
  F\"urther Strasse 248, 90429 N\"urnberg, Germany}%
 \affiliation{Department of Chemical and Biological Engineering and Department of Physics, Friedrich-Alexander-Universit\"at Erlangen-N\"urnberg, F\"{u}rther Stra{\ss}e 248, 90429 N\"{u}rnberg, Germany}
\date{\today}

\begin{abstract}
We study the dynamics of an unstable thin film deposited on a ``switchable`` substrate.
The substrate is itself subject to a time varying function that affects the equilibrium contact angle $\theta_{\text{eq}}(\mathbf{x},t)$.
We show that in the case of vanishing time dependence the patterning drives the dewetting morphology and that all fluid is drained into regions of small $\theta_{\text{eq}}(\mathbf{x})$ and thus form droplets.
Interestingly enough with a time varying component to the pattern we are able to stop the droplet fragmentation and create pronounced ligament or rivulet morphologies.
We show that both the numbers of semistable rivulets as well as number of stable droplets can be controlled with the pattern wavelength ($\lambda$).
Finally we explain why the ligaments appear with a simple balance between capillary retraction and dynamic wetting.   
\end{abstract}

@@

\section{Introduction}\label{sec:intro}
Thin film flows and especially droplets are an phenomenon we experience in our every day life.
Being it the droplets condensed in the higher atmosphere and return to the ground as rain or the oil film in a pan during cooking. 
While this two examples are illustrative there are many industrial applications that build on thin film flows. 
Among them is the wetting of surfaces to reduce wear or put simply, lubrication~\cite{ReynoldsLubr, gross1980fluid, szeri2010fluid}. 
Coating surfaces to change their physio-chemical properties is another widely used application~\cite{DERYCK1998278, doi:10.1146/annurev.fluid.31.1.347, DASILVASOBRINHO19991204}.
On the other hand the need to control droplets is at the heart of many printing processes~\cite{singh2010inkjet, jo2009evaluation}.

More recently lab-on-a-chip devices have acquired much attention~\cite{C6LC00387G,Focke}, to work they require precise droplet or film flow on small scales~\cite{stone2004engineering, darhuber2010planar}.
One possibility for precise control is the local wettability.
Light switchable substrates~\cite{WANG200718}, therefore substrates that change their wettability based on light illumination, open up a new way to program this devices. 
From a theoretical point of view the motion of the fluid can be computed using the thin film equation~\cite{RevModPhys.69.931, RevModPhys.81.1131, RevModPhys.81.739, ReynoldsLubr}
\begin{equation}\label{eq:thin_film}
    \partial_t h(\mathbf{x},t) = \nabla\cdot\left(M(h)\nabla p(\mathbf{x},t)\right),
\end{equation}
where $h(\mathbf{x},t)$ is a time and spatial varying film thickness, $M(h)$ is a mobility function and $p(\mathbf{x},t)$ is a pressure term that drives the flow. 
The function $M(h)$ contains information about the velocity boundary condition at $h=0$, using no-slip yields $M(h) = h^3/(3\mu)$.
More interesting however is the pressure function $p(\mathbf{x},t)$ as it allows to model the wettability of a given substrate.
We therefore introduce a time and spatial dependent contact angle to control the local wettability.

Spatial variations of the contact angle are long known and used in previous studies~\cite{PhysRevE.91.023013, Sommers_2006}, time dependency of the contact angle on the other hand is not as well studied~\cite{suman2006dynamics}. 
Switchable substrates are in fact a realization of a time dependent contact angle. 
Using both spatial as well as time dependent variations of the local contact angle leads to interesting phenomena, for example continuously sliding droplets as shown by Grawitter and Stark~\cite{D0SM02082F}.  
Our approach builds on the same idea but our method, see ref~\cite{PhysRevE.100.033313}, allows us to go beyond a single droplet and to simulate the full dewetting transition starting with a thin film and ending with droplets.
We use this to show that a time dependent component of the contact angle alter the dewetting morphology quite substantially.
In the limit of a small time dependency or slow evolution of the contact angle field we qualitatively reproduce the results from Grawitter and Stark, such sliding droplets.
Speeding up the evolution of the contact angle field introduces a morphological transition, which to the understanding of the authors have not been looked at so far.
Therefore the purpose of this manuscript is to increase the understanding of the dewetting process.
Highlighting the fact that a temporal and spatial variation of the contact angle leads to novel morphological structures.
Thus our finding offer a simple way to understand how time periodic switchable substrates could generate novel dewetting morphologies.

The manuscript is outlined as follows:
In section~\ref{sec:method} we discuss the numerical method we use to generate our results and introduce relevant scales.
We then present our results in section~\ref{sec:results}, where we discuss the impact of the time dependency on the contact angle and show that both stability as well as dewetting morphology is affected.
A short summary of our finding and a closing conclusion with outlook is given in section~\ref{sec:sum_conclu}.

\section{Method}\label{sec:method}
In this section we describe the lattice Boltzmann method (LBM) we use.
The LBM is a meso scale approach that builds on the Boltzmann equation to simulate fluid flows~\cite{krueger2017}. 
It is widely applied to many kinds of flow problems, among them are for instance multi phase and multi component flows~\cite{shan1993lattice,PhysRevE.73.047701}, flows through porous media~\cite{liu2016multiphase} and shallow water flows~\cite{Salmon:1999:0022-2402:503, zhou2004lattice, PhysRevE.65.036309}.
The model we use has been introduced in ref.~\cite{PhysRevE.100.033313} and builds upon a class of lattice Boltzmann models originally developed to study the shallow water equations~\cite{Salmon:1999:0022-2402:503, PhysRevE.65.036309}.

\subsection{Thin film lattice Boltzmann method}\label{subsec:LBM}
Since the model has been developed and tested in previous work, ref.~\cite{PhysRevE.100.033313}, we just outline the most important features.
The internal degrees of freedom of this method are the distributions functions $f_{\alpha}(\mathbf{x},t)$.
These distribution functions describe the movement of fictitious particles on a lattice, in this case on a $D2Q9$ lattice.
Meaning there are nine velocity components ($Q9$), directions in which a particle could move, for a simulation in two dimensions ($D2$).
The dynamics of the thin film and thus the evolution of the distribution functions is given by
\begin{equation}\label{eq:LBE}
    \begin{split}
        &f_{\alpha}(\mathbf{x}+\mathbf{c}^{({\alpha})}\Delta t,t+\Delta t) = \\
        &\left(1 - \frac{\Delta t}{\tau}\right) f_i(\mathbf{x},t) + \frac{\Delta t}{\tau} f_{\alpha}^{(eq)}(\mathbf{x},t) + \Delta t\mathcal{S}_{\alpha},
    \end{split}
\end{equation}
where the $\mathbf{c}^{({\alpha})}$ are the $Q$ discrete lattice velocities that can be computed according to
\begin{equation}\label{eq:speeds}
\mathbf{c}^{({\alpha})}  =
\left\{
\begin{array}{ll}
(0,0) & \alpha = 0 \\
\left[\cos{\frac{(\alpha-1)\pi}{4}}, \sin{\frac{(\alpha-1)\pi}{4}} \right] &  \alpha=1,3,5,7 \\
\sqrt{2}\left[\cos{\frac{(\alpha-1)\pi}{4}}, \sin{\frac{(\alpha-1)\pi}{4}} \right] & \alpha=2,4,6,8
\end{array}
\right.,
\end{equation}
and $w$ are being the corresponding weights
\begin{equation}
w_{\alpha}  =
\left\{
\begin{array}{ll}
\frac{4}{9} & \alpha = 0 \\
\frac{1}{9} &  \alpha=1,3,5,7 \\
\frac{1}{36} & \alpha=2,4,6,8
\end{array}
\right..
\end{equation}
The last term to the right $\mathcal{S}_{\alpha}$ is a (generalized) source term.
The concrete implementation of the forces depend on the source term as~\cite{https://doi.org/10.1002/fld.4726}
\begin{equation}\label{eq:source_term}
    \mathcal{S}_{\alpha} = \frac{3w_{\alpha}}{\mathbf{c}_{\alpha}^2}\mathbf{c}_{\alpha}\cdot\mathbf{F}.
\end{equation}
We are able to match thin film dynamics with the usage of a force term comprising of two main components,
\begin{equation}\label{eq:forcesum}
    \mathbf{F} = \mathbf{F}_{\mbox{\tiny{fric}}} + \mathbf{F}_{\mbox{\tiny{film}}}.
\end{equation}
The first being a friction force that includes a velocity boundary condition at the fluid substrate interface
\begin{equation}\label{eq:forcefric}
\mathbf{F}_{\mbox{\tiny{fric}}} = -\frac{6h\mu\mathbf{u}}{2h^2+6\delta h + 3\delta^2},
\end{equation}
where $\mu$ is the viscosity and $\delta$ is a socalled slip length~\cite{RevModPhys.69.931, PhysRevE.100.033313, PhysRevE.63.011208, munch_lubrication_2005}.
The film thickness $h$ is the zeroth moment of the distribution function $f_{\alpha}$, i.e. $h(\mathbf{x},t) = \sum_{\alpha}f_{\alpha}(\mathbf{x},t)$ see Eq.~(\ref{eq:LBE}). 
The second component is due to the film pressure $p_{\text{film}}$ and accounts for the minimization of the fluid vapor interface as well as for the wettability of the substrate,
\begin{equation}\label{eq:pressure_here}
    p_{\text{film}} = -\gamma (\Delta h - \Pi(h)),
\end{equation}
which results in the addition of the force $\mathbf{F}_{\mbox{\tiny{film}}}$
\begin{equation}\label{eq:forcefilm}
    \mathbf{F}_{\mbox{\tiny{film}}} = -\frac{h}{\rho_0}\nabla(p_{\text{film}}), 
\end{equation}
with $\gamma$ being the (constant) surface tension and $\rho_0$ being the fluids density, which is $\rho_0 = 1$ and will be dropped in the following. 
Here we model the interaction between solid substrate and film with a disjoining pressure term~\cite{RevModPhys.69.931, RevModPhys.81.1131, PhysRevE.63.011208},
\begin{equation}\label{eq:disjoinpressure}
    \Pi(h) = \kappa(\theta_{\text{eq.}})\left[\left(\frac{h_{\ast}}{h}\right)^n - \left(\frac{h_{\ast}}{h}\right)^m\right],
\end{equation}
with $n$ and $m$ being integer exponents, $h_{\ast}$ defines the minimum of the potential and  
\begin{equation}\label{eq:kappacontactangle}
    \kappa(\theta_{\text{eq.}}) = (1-\cos\theta_{\text{eq.}}) \frac{(n-1)(m-1)}{(n-m)h_{\ast}},
\end{equation}
which therefor allows us to tune the fluids contact angle.

@@

We consider two different patterns and three wavelengths in this manuscript.
First being the sinusoidal pattern with wavelength $\lambda = L/i$ and wavenumbers $i=[1,2,3]$ which is given as
\begin{equation}\label{eq:sinetheta}
    \theta_1(x,y) = \theta_0 + \delta\theta\left[\sin\left(\frac{2\pi x}{\lambda}\right)\sin\left(\frac{2\pi y}{\lambda}\right)\right], 
\end{equation}
with $\theta_0 = 20^{\circ}$ and $\delta\theta=10^{\circ}$.
The second pattern is a triangle wave which accounts for a different kind of pattern gradient.
To the best of our knowledge contact angles can not be tuned precise enough to match a sinusoidal wave on the scales we are interested in.
Therefore we assume that a triangle wave with a linear gradient is more likely accessable in experiments, thus $\theta_2$ is given as
\begin{equation}\label{eq:triangle_wave}
\begin{split}
    \theta_2(x,y) &= \theta_0 + \\ &\frac{2\delta\theta}{\pi}\left[\sin^{-1}\left(\sin\left(\frac{2\pi x}{\lambda}\right)\right)\sin^{-1}\left(\sin\left(\frac{2\pi y}{\lambda}\right)\right)\right].
    \end{split}
\end{equation}

Both patterns create a contact angle distribution with values within the interval $[10^{\circ}, 30^{\circ}]$ and by definition they are time independent.
To have a time dependency we simply shift the pattern in a periodic manner, such that
\begin{equation}\label{eq:theta_shift}
    \theta(\mathbf{x},t+T) = \theta(\mathbf{x}-\mathbf{1}, t), 
\end{equation}
with $T$ being an arbitrary integer number of lattice Boltzmann time steps.

Due to convenience we attribute a velocity to the pattern which is the shift distance divided by the stepping time
\begin{equation}\label{eq:pattern_speed}
    v_{\theta} = \frac{\sqrt{2}}{T}.
\end{equation}
Since this velocity as well as the choice of $T$ is arbitrary we define a characteristic velocity $v_0$ similar to Eqs.~(\ref{eq:q0_and_t0}) as
\begin{equation}\label{eq:normvel}
    v_0 = \frac{\lambda}{t_0(\theta)},
\end{equation}
where we use the wavelength $\lambda$ as a characteristic distance and $t_0(30^{\circ})$ as a characteristic time scale.
Important to note is that $t_0$ is sensitive to the contact angle, and decreases with increasing angle.

>>>>

\title{Switchable substrates and thin film morphologies}

\author{S. Zitz}
\email{s.zitz@fz-juelich.de}
 \affiliation{Helmholtz Institute Erlangen-N\"urnberg for Renewable Energy,\\
  Forschungszentrum J\"ulich,\\
  F\"urther Strasse 248, 90429 N\"urnberg, Germany}%
  \affiliation{Department of Chemical and Biological Engineering, Friedrich-Alexander-Universit\"at Erlangen-N\"urnberg, F\"{u}rther Stra{\ss}e 248, 90429 N\"{u}rnberg, Germany}
\author{A. Scagliarini}%
\email{andrea.scagliarini@cnr.it}
 \affiliation{Institute for Applied Mathematics "M. Picone" (IAc, \\
Consiglio Nazionale delle Ricerche,\\
Via dei Taurini 19, 00185 Rome, Italy}%
\affiliation{INFN, sezione Roma ``Tor Vergata'', via della Ricerca Scientifica 1, 00133 Rome, Italy}
\author{J. Harting}
\email{j.harting@fz-juelich.de}
 \affiliation{Helmholtz Institute Erlangen-N\"urnberg for Renewable Energy,\\
  Forschungszentrum J\"ulich,\\
  F\"urther Strasse 248, 90429 N\"urnberg, Germany}%
 \affiliation{Department of Chemical and Biological Engineering and Department of Physics, Friedrich-Alexander-Universit\"at Erlangen-N\"urnberg, F\"{u}rther Stra{\ss}e 248, 90429 N\"{u}rnberg, Germany}
\date{\today}

\begin{abstract}
We study the dynamics of an unstable thin film deposited on a switchable substrate.
The substrate is itself subject to a time varying function that affects the equilibrium contact angle $\theta_{\text{eq}}(\mathbf{x},t)$.
We show that in the case of the vanishing time dependence the patterning drives the dewetting process and all fluid is drained into regions of a small $\theta_{\text{eq}}(\mathbf{x})$ and forms droplets.
Interestingly enough with a time varying equilibrium contact angle we are able to stop the droplet fragmentation and are able to create ligament or rivulet states.
We show that these rivulet states mainly depend on two quantities, first being the wavelength of the pattern and second being the velocity of the substrate pattern.

\end{abstract}

@@

\section{Introduction}\label{sec:intro}
Thin film flows and especially droplets are an phenomenon we experience in our every day life.
Rather unknown is the fact that our body uses thin films to cover mucous membranes.

\begin{equation}\label{eq:thin_film}
    \partial_t h(\mathbf{x},t) = \nabla\cdot\left(M(h)\nabla p(\mathbf{x},t)\right),
\end{equation}

\section{Method}\label{sec:method}
In this section we describe the lattice Boltzmann method (LBM) we use to simulate thin film flows.
The model we use has been introduced in ref~\cite{PhysRevE.100.033313} and builds upon a class of lattice Boltzmann models originally developed to study the shallow water equations~\cite{Salmon:1999:0022-2402:503, zhou2004lattice, PhysRevE.65.036309}.

\subsection{Thin film lattice Boltzmann method}\label{subsec:LBM}
The internal degrees of freedom of this method are the distributions functions $f_{\alpha}(\mathbf{x},t)$.
They describe a distribution of fictitious particles, for this paper we consider the case of $D2Q9$ lattice.
Such there are nine velocity components ($Q9$) for a simulation in two dimensions ($D2$).
The dynamics of the thin film and thus the evolution of the distribution functions is given by
\begin{equation}\label{eq:LBE}
    \begin{split}
        &f_{\alpha}(\mathbf{x}+\mathbf{c}^{({\alpha})}\Delta t,t+\Delta t) = \\
        &\left(1 - \frac{\Delta t}{\tau}\right) f_i(\mathbf{x},t) + \frac{\Delta t}{\tau} f_{\alpha}^{(eq)}(\mathbf{x},t) + \Delta t\mathcal{S}_{\alpha},
    \end{split}
\end{equation}
where the $\mathbf{c}^{({\alpha})}$ are the $Q$ discrete lattice velocities that can be computed according to
\begin{equation}\label{eq:speeds}
\mathbf{c}^{({\alpha})}  =
\left\{
\begin{array}{ll}
(0,0) & \alpha = 0 \\
\left[\cos{\frac{(\alpha-1)\pi}{4}}, \sin{\frac{(\alpha-1)\pi}{4}} \right] &  \alpha=1,3,5,7 \\
\sqrt{2}\left[\cos{\frac{(\alpha-1)\pi}{4}}, \sin{\frac{(\alpha-1)\pi}{4}} \right] & \alpha=2,4,6,8
\end{array}
\right.,
\end{equation}
and $w$ are being the corresponding weights
\begin{equation}
w_{\alpha}  =
\left\{
\begin{array}{ll}
\frac{4}{9} & \alpha = 0 \\
\frac{1}{9} &  \alpha=1,3,5,7 \\
\frac{1}{36} & \alpha=2,4,6,8
\end{array}
\right..
\end{equation}
The last term to the right $\mathcal{S}_{\alpha}$ is a (generalized) source term.
The concrete implementation of the forces depend on the source term as~\cite{https://doi.org/10.1002/fld.4726}
\begin{equation}\label{eq:source_term}
    \mathcal{S}_{\alpha} = \frac{3w_{\alpha}}{\mathbf{c}_{\alpha}^2}\mathbf{c}_{\alpha}\cdot\mathbf{F}.
\end{equation}
We are able to match thin film dynamics with the usage of a force term comprising of two main components,
\begin{equation}\label{eq:forcesum}
    \mathbf{F} = \mathbf{F}_{\mbox{\tiny{fric}}} + \mathbf{F}_{\mbox{\tiny{fric}}}.
\end{equation}
The first being a friction force that includes a velocity boundary condition at the fluid substrate interface
\begin{equation}\label{eq:forcefric}
\mathbf{F}_{\mbox{\tiny{fric}}} = -\frac{6h\mu\mathbf{u}}{2h^2+6\delta h + 3\delta^2},
\end{equation}
where $\mu$ is the viscosity and $\delta$ is a socalled slip length~\cite{RevModPhys.69.931, PhysRevE.100.033313, PhysRevE.63.011208, munch_lubrication_2005}.
The film thickness $h$ is the zeroth moment of the distribution function $f_{\alpha}$, i.e. $h(\mathbf{x},t) = \sum_{\alpha}f_{\alpha}(\mathbf{x},t)$ see Eq.~(\ref{eq:LBE}). 
The second component is due to the film pressure and accounts for the minimization of the fluid vapor interface as well as for the wettability of the substrate,
\begin{equation}\label{eq:forcefilm}
    \mathbf{F}_{\mbox{\tiny{film}}} = -\frac{\gamma h}{\rho_0}\nabla(\Delta h - \Pi(h)), 
\end{equation}
with $\gamma$ being the surface tension and $\rho_0$ being the fluids density, which is $\rho_0 = 1$ and will be dropped in the following. 
Here we model the interaction between solid substrate and film with a disjoining pressure term~\cite{RevModPhys.69.931, RevModPhys.81.1131, PhysRevE.63.011208},
\begin{equation}\label{eq:disjoinpressure}
    \Pi(h) = \kappa(\theta_{\text{eq.}})\left[\left(\frac{h_{\ast}}{h}\right)^n - \left(\frac{h_{\ast}}{h}\right)^m\right],
\end{equation}
with $n$ and $m$ being integer exponents, $h_{\ast}$ defines the minimum of the potential and  
\begin{equation}\label{eq:kappacontactangle}
    \kappa(\theta_{\text{eq.}}) = (1-\cos\theta_{\text{eq.}}) \frac{(n-1)(m-1)}{(n-m)h_{\ast}},
\end{equation}
which therefor allows us to tune the fluids contact angle.

@@

We consider two different patterns and three wavelengths for this manuscript.
First being the sinusoidal pattern with wavelengths $\lambda \in \{L, L/2, L/3\}$ which is given as
\begin{equation}\label{eq:sinetheta}
    \theta_1(x,y) = \theta_0 + \delta\theta\left[\sin\left(\frac{2\pi x}{\lambda}\right)\sin\left(\frac{2\pi y}{\lambda}\right)\right], 
\end{equation}
with $\theta_0 = 20^{\circ}$ and $\delta\theta=10^{\circ}$.
The second pattern is a triangle wave which accounts for a more realistic szenario, thus $\theta_2$ is given as
\begin{equation}\label{eq:triangle_wave}
    \theta_2(x,y) = \theta_0 + \frac{2\delta\theta}{\pi}\left[\sin^{-1}\left(\sin\left(\frac{2\pi x}{\lambda}\right)\right)\sin^{-1}\left(\sin\left(\frac{2\pi y}{\lambda}\right)\right)\right].
\end{equation}

Both patterns create a contact angle distribution with values within the interval $[10^{\circ}, 30^{\circ}]$ and by definition they are time independent.
For a time dependent dynamic we simply shift the pattern in a periodic manner, such that
\begin{equation}\label{eq:theta_shift}
    \theta(\mathbf{x},t+T) = \theta(\mathbf{x}-\mathbf{1}, t), 
\end{equation}
with $T$ being an arbitrary number of lattice Boltzmann time steps.

Due to convenience we artibute a velocity to the pattern which is simply the shift distance divided by the stepping time
\begin{equation}\label{eq:pattern_speed}
    v_{\theta} = \frac{\sqrt{2}}{T}.
\end{equation}
Since this velocity as well as the choice of $T$ is arbitrary we define a characteristic velocity $v_0$ similar to Eqs.~(\ref{eq:q0_and_t0}) as
\begin{equation}\label{eq:normvel}
    v_0 = \frac{\lambda}{t_0(\theta)},
\end{equation}
where we use the wavelength $\lambda$ as a characteristic distance and $t_0(30^{\circ})$ as a characteristic time scale.
Important to note is that $t_0$ is sensitive to the contact angle, and decreases with increasing angle.

\subsection{Numerical system}
As outlined in Sec.~\ref{subsec:LBM} we simulate the evolution of the thin film with a lattice Boltzmann method.
For all our simulations we use periodic boundary conditions and a square of size $Lx = Ly = L = 512 \Delta x$ with $\Delta x = \Delta t = 1$.
The surface tension as well as the viscosity are constants and set to $\gamma = 0.01$ and $\mu = 1/6$.
Having this value of $\mu$ leaves us with a relaxation time $\tau = 1$.
To regularize the contact line divergence~\cite{huh1971hydrodynamic} we use a precursor layer and a slip length $\delta = 1$. 
With this choice of parameters we are almost able simulate a no-slip boundary condition.
As for our disjoining pressure potential, we use $n=3, m=2$ and $h_{\ast} = 0.07$.
Inserting this quantities into Eq.~(\ref{eq:q0_and_t0}) with $h_0 = 1$ yields $q_0^2(30^{\circ}) \approx 1.68\cdot10^{-2}$ as well as $t_0(30^{\circ}) \approx 177000\Delta t$ which gives us a first glimpse into relevant time scales.
